---
title: "Homework 6"
author: "Madison Goldrich"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Load key packages.

```{r load_libraries}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New York. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

The code below (provided with the assignment instructions) downloads the Central Park weather data.

```{r q2_data_import}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

The below code chunk creates a bootstrap sampling function, then applies that function create 5000 bootstrap samples for our `weather_df`.

```{r q2_bootstrap_samples}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

boot_straps =
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df))
  )
```

The below code chunk produces estimates of $\hat{r}^2$ and $\log(\hat{\beta}_1 * \hat{\beta}_2)$ for each sample.

```{r q2_boot_analysis}
boot_tidy_results =
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    tidy_results = map(models, broom::tidy),
  ) |> 
  select(strap_number, tidy_results) |> 
  unnest(tidy_results) |>
  select(strap_number, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(log_beta1_beta2 = log(tmin * prcp)) |> 
  select(strap_number, log_beta1_beta2)

boot_glance_results =
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    glance_results = map(models, broom::glance),
  ) |> 
  select(strap_number, glance_results) |> 
  unnest(glance_results) |>
  select(strap_number, r_squared = r.squared)

boot_results =
  inner_join(boot_tidy_results, boot_glance_results)
```

The below code chunk produces two histograms to visualize the distribution of our two estimates. We can see that $\hat{r}^2$ is very close to normally distributed. There is a very slight left-tail. The values appear to range from approximately 0.865 to 0.95. From the second histogram, we can see that $\log(\hat{\beta}_1 * \hat{\beta}_2)$ has a much more distinct left-tailed skew. Further, a number of rows indicated by the warning message had to be removed for containing non-finite values (caused by attempting to take the log of a negative value).

```{r q2_plots, warning = TRUE}
boot_results |> 
  ggplot(aes(x = r_squared)) + 
  geom_histogram() +
  labs(title = "R_squared distribution")

boot_results |> 
  ggplot(aes(x = log_beta1_beta2)) + 
  geom_histogram() +
  labs(title = "Log(beta1 * beta2) distribution")
```

The below code chunk produces the 95% confidence intervals for $\hat{r}^2$ and $\log(\hat{\beta}_1 * \hat{\beta}_2)$. We are 95% confident that the true $\hat{r}^2$ lies between 0.889 and 0.941, and we are 95% confident that the true $\log(\hat{\beta}_1 * \hat{\beta}_2)$ lies between -8.963 and -4.603.

```{r q2_CIs}
boot_results |> 
  summarize(
    r_squared_CI_lower = quantile(r_squared, 0.025),
    r_squared_CI_upper = quantile(r_squared, 0.975),
    log_beta1_beta2_CI_lower = quantile(log_beta1_beta2, 0.025, na.rm = TRUE),
    log_beta1_beta2_CI_upper = quantile(log_beta1_beta2, 0.975, na.rm = TRUE)
  ) |> 
  knitr::kable(digits = 3)
```

### Problem 3

This code chunk imports the csv file and cleans the data for regression analysis, namely converting the categorical variables from numeric to factor format.

```{r q3_dataimport}
birthweight_df =
  read_csv("data/birthweight.csv") |> 
  mutate(
    babysex = fct_infreq(as.factor(babysex)),
    frace = fct_infreq(as.factor(frace)),
    malform = fct_infreq(as.factor(malform)),
    mrace = fct_infreq(as.factor(mrace)))
```

The following code chunk performs linear regression of this model: `bwt = blength + gaweeks + babysex + delwt + momage`. I chose this model based on the other two models that I will be comparing it to, which state that length at birth and gestational age are the main effects. I will also be looking at sex and its interaction with the other two effects, so I added in sex. Finally, I thought that mother's weight and age at delivery would be interesting to include. The table produced gives the estimate and p-value for each term.

```{r q3_regression}
fit = 
  lm(
    bwt ~ blength + gaweeks + babysex + delwt + momage,
    data = birthweight_df)

fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(term = str_replace(term, "babysex2", "baby sex: female")) |> 
  knitr::kable()
```

Below is a code chunk that produces a scatterplot of model residuals against fitted values.

```{r q3_plotresiduals}
birthweight_df |> 
  modelr::add_predictions(fit) |> 
  modelr::add_residuals(fit) |> 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point(alpha = .3) +
  labs(title = "Model residuals vs. fitted values")
```







